Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations

### 2021-12-22 analysis/paper/paper.Rmd
draft Results complete

@@

@@

@@

To estimate the total consumption potential (i.e., the number or amount of juvenile Chinook Salmon consumed during the peaks of DSR [fall] and NRR [spring] emigrations) of Northern Pikeminnow in Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 and applied in the R statistical software [@RCoreTeam2021]. The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on the following inputs: predator and prey energy densities, predator start and end weights, and water temperatures. Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017]. Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon [@Moss2016]. The predator start weight was calculated using a weight-length formula [@Parker1995] and the average length of Northern Pikeminnow caught in Deadwater Slough during our study. The average length of Northern Pikeminnow caught during the fall and spring efforts were 352.9 and 393.7 mm, respectively, which calculated to average starting weights of 504.4 and 670.5 g. We assumed no growth in individual Northern Pikeminnow, and so predator start and end weights are equal, i.e., we assumed the population biomass was stable. Mean daily water temperatures were summarized from 15-minute interval temperature readings available March 3, 2013 to June 14, 2021 from a gage station approximately 22 river kilometers downstream of Deadwater Slough near Shoup, Idaho [@USGS2016].

@@

We estimated a single average-sized Northern Pikeminnow to consume xx.x g of fish during the modeled fall time period (September 15 - November 30) and an additional xx.x g of fish during the modeled spring time period (March 1 - June 31). Given estimated population sizes of xx,xxx Northern Pikeminnow during fall and xx,xxx during spring, that would convert to xxx,xxx - xxx,xxx g during fall and xxx,xxx g during spring. Given average weights of 10.3 g for DSR Chinook salmon and 10.9 g for NRR Chinook salmon, we would estimate that xx,xxx and xx,xxx juvenile Chinook salmon could be consumed during the fall and spring emigrations in Deadwater Slough, respectively.

The estimated number of Chinook Salmon consumed by Northern Pikeminnow begins to taper off between 30-90% of Pikeminnow diet. So although we are uncertain of that diet proportion, the impact of that uncertainty is somewhat constrained. We estimated the total juvenile outmigrants consumed (fall and spring) to be between `r paste(prettyNum(round(range(total_impacts$smolt_N)), big.mark = ","), collapse = " and ")`. This translates to adult Chinook equivalents between `r paste(prettyNum(round(range(total_impacts$adult_N)), big.mark = ","), collapse = " and ")` (Figure \@ref(fig:adult-imp-p)).

@@

>>>>

# average size of fall and spring pikeminnow
avg_fall_NPM   = 352.9
avg_spring_NPM = 393.7

@@

fall_g_summ = fall_results %>%
  select(f10_g:f90_g) %>%
  pivot_longer(f10_g:f90_g) %>%
  group_by(name) %>%
  summarize(cumm_g = max(value))

@@

spring_g_summ = spring_results %>%
  select(f10_g:f90_g) %>%
  pivot_longer(f10_g:f90_g) %>%
  group_by(name) %>%
  summarize(cumm_g = max(value))

@@

To estimate the total consumption potential (i.e., the number or amount of juvenile Chinook Salmon consumed during the peaks of DSR [fall] and NRR [spring] emigrations) of Northern Pikeminnow in Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 and applied in the R statistical software [@RCoreTeam2021]. The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on the following inputs: predator and prey energy densities, predator start and end weights, and water temperatures. Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017]. Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon [@Moss2016]. The predator start weight was calculated using a weight-length formula [@Parker1995] and the average length of Northern Pikeminnow caught in Deadwater Slough during our study. The average length of Northern Pikeminnow caught during the fall and spring efforts were `r avg_fall_NPM` and `r avg_spring_NPM` mm, respectively, which calculated to average starting weights of 504.4 and 670.5 g. We assumed no growth in individual Northern Pikeminnow, and so predator start and end weights are equal, i.e., we assumed the population biomass was stable. Mean daily water temperatures were summarized from 15-minute interval temperature readings available March 3, 2013 to June 14, 2021 from a gage station approximately 22 river kilometers downstream of Deadwater Slough near Shoup, Idaho [@USGS2016].

@@

We estimated an individual, average-size (`r avg_fall_NPM` mm) Northern Pikeminnow to consume `r median(fall_g_summ$cumm_g)` g (range `r min(fall_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the DSR emigration, depending on the proportion of diet consisting of fish versus invertebrates (Figure \@ref(fig:bio-p)). Similarly, we estimated the average-size (`r avg_spring_NPM` mm) Northern Pikeminnow to consume `r median(spring_g_summ$cumm_g)` g (range `r min(spring_g_summ$cumm_g)` - `r max(fall_g_summ$cumm_g)` g) of fish during the NRR emigration (Figure \@ref(fig:bio-p)). Given an average weight of `r avg_DSR_g` g for DSR emigrants, that would equate to `r round(median(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` (range `r round(min(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Fall"]), 2)`) fish consumed by an individual during fall and an additional `r round(median(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` (range `r round(min(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)` - `r round(max(adult_impact$juv_chnk[adult_impact$season == "Spring"]), 2)`) fish consumed using an average weight of `r avg_NRR_g` g for NRR emigrants.

Given a mean population size of `r prettyNum(unique(adult_impact$N[adult_impact$season == "Fall"]), big.mark = ",")` from fall estimates and an estimate of `r prettyNum(round(unique(adult_impact$N[adult_impact$season == "Spring"]), 0), big.mark = ",")` for spring 2021, and assuming the median value from diet scenarios (i.e, 50% of their diet consists of fish), we'd estimate that `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Fall"]), 0), big.mark = ",")`) DSR and `r prettyNum(round(median(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` (95% CI: `r prettyNum(round(min(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")` - `r prettyNum(round(max(adult_impact$smolt_N[adult_impact$season == "Spring"]), 0), big.mark = ",")`) NRR juveniles may be consumed in Deadwater Slough during peak emigrations. Table \@ref(tab:adult-imp) shows the estimated number of juvenile Chinook Salmon consumed by and individual Northern Pikeminnow and by the Deadwater Slough population across all diet scenarios and for both the fall (DSR) and spring (NRR) peak emigration seasons.

Across both seasons and all diet scenarios, we estimated the total juvenile outmigrants consumed to be between `r prettyNum(round(min(total_impacts$smolt_N), 0), big.mark = ",")` and `r prettyNum(round(max(total_impacts$smolt_N), 0), big.mark = ",")`. The estimated number of Chinook Salmon consumed by a Northern Pikeminnow begins to taper off betwen the 30-90% diet scenarios. So although we are uncertain of the diet proportions, the impact of that uncertainty is somewhat constrained.

@@

This translates to adult Chinook equivalents between `r paste(prettyNum(round(range(total_impacts$adult_N)), big.mark = ","), collapse = " and ")` (Figure \@ref(fig:adult-imp-p)).
